---
title: "London Fire Brigade Calls EDA - Analysis"
author: "Cem Gürkan - BDA 503 - Fall 2017"
output: 
  html_document: 
    highlight: zenburn
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is my first data analysis on R. Data is related with Calls made to London Fire Brigade between Jan 01 2017 to April 30 2017. Dataset is taken from [Kaggle|London Fire Data](https://www.kaggle.com/jboysen/london-fire/data).

I'll do preliminary analysis on data and answer some questions and some visualizations.

Hope you enjoy :)

## Loading Data and Preliminary Analysis

Load libraries used for this study.

```{r warning=FALSE, message=FALSE}
# Load packages
library('ggplot2') # visualization
library('ggthemes') # visualization
library('dplyr') # data manipulation
library('scales') # visualization
```

Let's import data set without header line.

```{r include=TRUE, tidy=TRUE}

#Read the data and omit the first row
calls = read.csv("~/london_fire_brigade_service_calls.csv", header = TRUE)

#See the dimension of our data
dim(calls)
```

Our data consisting of 32 columns and 32247 rows. Let's what are these columns and their contents

```{r include=TRUE, tidy=TRUE}
#Structure and sample contents
glimpse(calls)

# Summary of data
summary(calls) 
```

## Data Exploration and Visualization

By creating trend analysis we can see number of calls on each day.   

1. As you can see there is a outlier in Feb 23 and 523 calls made. 
1. Average 268 calls made

```{r}
# Find the numbers of call on each day.
grp_call <- calls %>%
  group_by(date_of_call) %>%
  summarise(count=n()) 

#Show on the graph, see the trends
ggplot(data=grp_call, aes(x=date_of_call, y=count, group = 1)) +
  geom_line(colour = "red") +
  labs(title="Trend of Calls between Jan-April 2017", x="Date of Call", y="Count") +
  theme (axis.text.x=element_text (angle=-90,vjust=0.5,hjust=0))

#Average calls
grp_call %>% summarise(count=mean(count)) 
```

If we examine calls on day,

* The number of calls increases from 0500 and reaches its peak at 1800
* Early mornings have lower calls

```{r}

ggplot(data=calls,aes(x=calls$hour_of_call, group=1)) +
geom_bar() +
labs(title="Call trend by the hour of day",x="Hour of Call",y="Count")

``` 

If we examine the response time;  
* Regardless of the hour of call, the average first response time is around 5 mins for all hours of the day

```{r}
data <- calls %>% mutate(first_pup_arriving_min=first_pump_arriving_attendance_time/60)

ggplot(data=data,aes(y=data$first_pup_arriving_min, x=data$hour_of_call)) +
  geom_boxplot(aes(group = cut_width(hour_of_call, 0.5)), outlier.alpha = 0.1, outlier.colour = "red", colour = "#3366FF")+
labs(title="Average First Response Time",x="Hour of Call",y="Minute")
``` 

What about if we examine the highest calls day  

* Average first response time is higher than all others day  

```{r}
#Highest call of the day
data <- calls %>% filter(date_of_call=="2017-02-23")
data <- data %>% mutate(first_pup_arriving_min=first_pump_arriving_attendance_time/60)


ggplot(data=data,aes(y=data$first_pup_arriving_min, x=data$hour_of_call)) +
  geom_boxplot(aes(group = cut_width(hour_of_call, 0.5)), outlier.alpha = 0.1, outlier.colour = "red", colour = "#3366FF")+
labs(title="Average First Response Time",x="Hour of Call",y="Minute")
``` 

If we want to look into which property type has most calls, Residential house is first one as expected. 


```{r}
#Find the #of calls for each property category
data <- calls %>%
  group_by(property_category) %>%
  summarise(count=n())  %>%
  arrange(desc(count))

data

ggplot(data=data,aes(x=reorder(property_category,-count), y=data$count)) +
  geom_bar(stat="identity") +
  labs(title="Calls for Property Category",x="Property Category",y="Count") +
  theme (axis.text.x=element_text (angle=90,vjust=0.5, hjust=0))


``` 
