---
title: 'Tidyverse Recap: Travel and Weather - Part I dplyr'
author: "BDA 503 - Fall 2017"
date: "20 Kasım 2017"
output:
  html_document:
    highlight: pygments
    theme: lumen
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This exercise is designed a recap to introduction to tidyverse from the very basics. 

Solution of the exercices in [this documents](https://mef-bda503.github.io/files/tidyverse_recap_p1.html) shown below.


## Preparation

First we are going to download data then load the weather data.

```{r cars}
# Load the package to the session
library(tidyverse)
# Download file from gitHub
download.file("https://mef-bda503.github.io/files/travel_weather.RData", "travel_weather.RData")
# Load the data set file
load("travel_weather.RData")
```

Now let’s take a look at our data.

## Travel Weather Data

```{r}
travel_weather %>%
    tbl_df()
```

## Final Exercises

1. Return the dates which Amsterdam is strictly warmer than London but strictly colder than Venice

```{r}
travel_weather %>% filter( Amsterdam>London & Amsterdam < Venice)
```


2. For each month of each year calculate the average difference between NYC and Amsterdam for the days NYC is strictly warmer than Amsterdam, rounded by 1 decimal. Arrange from the highest difference to the lowest.

```{r}
travel_weather %>% 
  filter(NYC>Amsterdam) %>%
  group_by (year, month) %>%
  summarise(NYCwA_diff = round(mean(NYC-Amsterdam),digits=1) ) %>%
  arrange(desc(NYCwA_diff))
```


3. Return the warmest city and its temperature of each day.

```{r}
travel_weather %>%
  gather(key=City,value=Temperature, -year, -month, -day) %>%
  group_by(year, month, day) %>%
  mutate(max_temp =max(Temperature)) %>%
  filter(Temperature==max_temp)  %>%
  arrange(year,month,day) %>%
  select(year, month, day,City,Temperature)
```